---
title: "FE570 - Homework #3"
subtitle: "I pledge my honor that I have abided by the Stevens Honor System."
author: "Sid Bhatia"
date: "2023-10-23"
output: pdf_document
---

Use the same dataset as in Assignment #2 and keep only the trades taking place in the time period 10:00 to 14:00.

```{r}
# Load necessary packages.
library(xts)
library(highfrequency)

# Load in data set.
options(digits.secs=3)
absolute_path <- 'C:/Users/sbhatia2/My Drive/University/Academics/Semester V/FE570 - Market Microstructure and Trading Strategies/FE570 - Homework/FE570 - Homework #2/'
load(paste(absolute_path, "sampleTQdata.RData", sep = ""))

# Added to remove warnings about time zone mismatch.
Sys.setenv(TZ='GMT')

head(tqdata)
tail(tqdata)

# Filter data for trades between 10:00 and 14:00
tqdata_filtered <- tqdata["T10:00/T14:00"]

head(tqdata_filtered)
tail(tqdata_filtered)
```

## Problem 1

1. How many trades are taking place during this time?

```{r}
num_of_trades <- nrow(tqdata_filtered)

num_of_trades
```

There are 4691 trades taking place during this time.

2. Compute the Quoted Spread and Effective Spread, averaged over all trades.

$$\text{Quoted Spread} = \text{Best Ask} - \text{Best Bid}$$

$$\text{Effective Spread} = 2 * D_t * (\text{Price}_t - \frac{1]{2}(a_t + b_t))$$

```{r}
colnames(tqdata_filtered)

# Retrieve liquidity measures using `highfrequency` package.
liqudity_measures <- getLiquidityMeasures(tqdata_filtered)

head(liqudity_measures$effectiveSpread)

# Retrieve trade signs using Lee-Ready method.
trade_signs <- getTradeDirection(tqdata_filtered)

price <- as.numeric(tqdata_filtered$PRICE)
mids <- (as.numeric(tqdata_filtered$OFR) + as.numeric(tqdata_filtered$BID)) / 2

# Compute Quoted Spread.
quoted_spread <- mean(as.numeric(tqdata_filtered$OFR) - as.numeric(tqdata_filtered$BID))
quoted_spread

# Compute Effective Spread.
effective_spread <- mean(trade_signs * (price - mids))
effective_spread    
```

3. Calibrate the Roll model to this data, and find the parameters $c$ (cost of trading) and $\sigma_u$ (volatility of the efficient price).

```{r}
# Calibrate Roll Model to data.

dprice <- diff(price)
ac_pr <- acf(dprice, lag.max=20, type="correlation", plot=FALSE)
plot(ac_pr, col="red", main="Autocorrelation of Price Changes")

covpr <- acf(dprice, lag.max=20, type="covariance", plot=FALSE)

gamma0 <- sd(dprice)^2
gamma0

gamma1 <- cov_prices$acf[2]
gamma1

cparam <- sqrt(-covpr$acf[2])
cparam

sig2u <- gamma0 + 2 * gamma1
sigu <- sqrt(sig2u)

cparam
sigu
```

As such, $c = 0.0158516$ and $\sigma_u =  0.04473154$.

4. Compute the Roll model estimate for the bid-ask spread, and compare with the spreads computed in point 2. Which one do you think is more accurate?

```{r}

```