---
title: "FE570 - Midterm Exam"
subtitle: "I pledge my honor that I have abided by the Stevens Honor System."
author: "Sid Bhatia"
date: "2023-10-23"
output: pdf_document
---

## Problem 11

The data for this problem is contained in the file *taqdata BTCUSD.RData*. This is a trade-and-quote file giving the trade price, size, and the quotes at
the time of each trade for Bitcoin trades during 24 hours (19-Apr-2023).

```{r}
# Load necessary packages.
library(xts)
library(highfrequency)

# Load in data set.
options(digits.secs=3)
absolute_path <- 'C:/Users/sbhatia2/My Drive/University/Academics/Semester V/FE570 - Market Microstructure and Trading Strategies/FE570 - Exams/FE570 - Midterm Exam/'
load(paste(absolute_path, "taqdata_BTCUSD.RData", sep = ""))

# Added to remove warnings about time zone mismatch.
Sys.setenv(TZ='GMT')

head(tqdata, 10)
```

#### i.

Report the number of trades in the dataset, and the minimum and maximum trade price during the time interval in the dataset.

```{r}
# Retrieve the number of trades in the dataset.
num_of_trades <- nrow(tqdata)
num_of_trades

price <- as.numeric(tqdata$PRICE)

# Establish minimum and maximum prices quoted.
p_min <- min(price)
p_max <- max(price)

p_min
p_max
```

#### ii.

For each transaction, compute the spread measures:

$$\text{Quoted Spread: qs_t = \text{Ask}_t - \text{Bid}_t}$$

$$\text{Effective Spread: es_t = 2d_t(p_t - \text{mid_t})}$$.

```{r}
# Compute the bids and asks for each transaction.
ask <- as.numeric(tqdata$OFR)
bid <- as.numeric(tqdata$BID)

# Compute the quoted spread.
quoted_spread <- ask - bid

head(quoted_spread, 50)
tail(quoted_spread, 50)

# Compute the mid prices (average of best bid and best ask prices).
mid <- (ask + bid) * 0.5

# Retrieve the trade sign for each transaction.
sign <- tqdata$SIDE

# Convert the trade sign for a "Buy" and "Sell" to 1 and -1, respectively.
sign_converted <- sign
sign_converted[sign_converted == "Buy"] <- 1
sign_converted[sign_converted == "Sell"] <- -1

sign_converted <- as.numeric(sign_converted)

head(sign, 10)
head(sign_converted, 10)

# Calculate the effective spread.
effective_spread <- 2 * sign_converted * (price - mid)

head(effective_spread, 50)
tail(effective_spread, 50)

mean(quoted_spread)
mean(effective_spread)
```