---
title: "FE570 - Midterm Exam"
subtitle: "I pledge my honor that I have abided by the Stevens Honor System."
author: "Sid Bhatia"
date: "2023-10-23"
output: pdf_document
---

## Problem 11

The data for this problem is contained in the file *taqdata BTCUSD.RData*. This is a trade-and-quote file giving the trade price, size, and the quotes at the time of each trade for Bitcoin trades during 24 hours (19-Apr-2023).

```{r}
# Load necessary packages.
library(xts)
library(highfrequency)

# Load in data set.
options(digits.secs=3)
absolute_path <- 'C:/Users/sbhatia2/My Drive/University/Academics/Semester V/FE570 - Market Microstructure and Trading Strategies/FE570 - Exams/FE570 - Midterm Exam/'
load(paste(absolute_path, "taqdata_BTCUSD.RData", sep = ""))

# Added to remove warnings about time zone mismatch.
Sys.setenv(TZ='GMT')

head(tqdata, 10)
```

#### i.

Report the number of trades in the dataset, and the minimum and maximum trade price during the time interval in the dataset.

```{r}
# Retrieve the number of trades in the dataset.
num_of_trades <- nrow(tqdata)
num_of_trades

price <- as.numeric(tqdata$PRICE)

# Establish minimum and maximum prices quoted.
p_min <- min(price)
p_max <- max(price)

p_min
p_max
```

The number of trades is $\mathbf{58793}$ with the minimum price at $\mathbf{28534.75}$ and maximum price at $\mathbf{30407.50}$.

#### ii.

For each transaction, compute the spread measures:

$$\text{Quoted Spread}: qs_t = \text{Ask}_t - \text{Bid}_t$$

$$\text{Effective Spread}: es_t = 2d_t(p_t - \text{Mid}_t)$$.

```{r}
# Compute the bids and asks for each transaction.
ask <- as.numeric(tqdata$OFR)
bid <- as.numeric(tqdata$BID)

# Compute the quoted spread.
quoted_spread <- ask - bid

head(quoted_spread, 50)
tail(quoted_spread, 50)

# Compute the mid prices (average of best bid and best ask prices).
mid <- (ask + bid) * 0.5

# Retrieve the trade sign for each transaction.
sign <- tqdata$SIDE

# Convert the trade sign for a "Buy" and "Sell" to 1 and -1, respectively.
sign_converted <- sign
sign_converted[sign_converted == "Buy"] <- 1
sign_converted[sign_converted == "Sell"] <- -1

sign_converted <- as.numeric(sign_converted)

head(sign, 10)
head(sign_converted, 10)

# Calculate the effective spread.
effective_spread <- 2 * sign_converted * (price - mid)

head(effective_spread, 50)
tail(effective_spread, 50)

mean(quoted_spread)
mean(effective_spread)
```

The average quoted spread is $\mathbf{4.817}$ and the average effective spread is $\mathbf{1.957}$.

#### iii.

Compute the Roll's estimate of the bid-ask spread.

```{r}
# Calculate the difference in price changes.
dprice <- diff(price)

# Compute and plot the autocorrelation of price changes.
ac_pr <- acf(dprice, lag.max=20, type="correlation", plot=FALSE)
plot(ac_pr, col="red", main="Autocorrelation of Price Changes")

# Compute the covariances of the price changes.
covpr <- acf(dprice, lag.max=20, type="covariance", plot=FALSE)

# Retrieve gamma1 as the covariance at lag 0.
gamma0 <- covpr$acf[1]
gamma0

# Retrieve gamma1 as the covariance at lag 1.
gamma1 <- covpr$acf[2]
gamma1

# Compute the volatility of the efficient price.
sig2u <- gamma0 + 2 * gamma1
sigu <- sqrt(sig2u)
sigu

# Compute the c paramater as the sqrt(-gamma1)
cparam <- sqrt(-gamma1)
cparam

# Compute the spread as 2 * the c parameter.
roll_spread <- cparam * 2

roll_spread
```

As such, the Roll's model estimate of the bid-ask spread is $\mathbf{4.756}$ with $c = 2.378$ and $\sigma_u = 3.255$.

#### iv.

Compare the trade sign in SIDE with the prediction of the Lee-Ready empirical rule. 

What is the accuracy of the Lee-Ready rule? 

This can be measured as the percentage of trade signs which are predicted correctly by the Lee-Ready rule.

**Tick Test**: Use only the trade prices $p_t$, but not the quotes $a_t$ and $b_t$. Under the test, the trade is classified as a buy/sell according to:
- $d_t = +1$ (buy) if $p_t > p_{t-1}$ (uptick) or if $p_t = p_{t-1} > p_{t-2}$ (zero-uptick)
- $d_t = -1$ (sell) if $p_t < p_{t-1}$ (downtick) or if $p_t = p_{t-1} < p_{t-2}$ (zero-downtick)

Note that zero-uptick/downtick results apply also if there are multiple (more than 2) trades with the same price.

For example if the trade prices are $p_t = (19.9, 20.0, 20.0, 20.0)$ (increasing $t$ order), then the trade signs are (?, +, + , +).

**Lee-Ready Rule**: Use both $p_t$ and quotes $a_t$ and $b_t$. The Lee-Ready Rule decides if a trade is a buy or sell by comparing the trade
price $p_t$ with the mid-price $m_t = \frac{1}{2} (a_t + b_t)$ (the half-point between best-bid $b_t$ and best-ask $a_t$).

If the trade price is exactly equal to the mid-price, $p_t = m_t$, then use the tick rule in point (i) above.

```{r}
# Create a function that implements the Tick Test.
tick_test <- function(price)
{
    sign <- c(1)
    for(i in 2:(length(price))) 
    {
        if(price[i] < price[i - 1])
        {
            sign <- c(sign, -1)
        }
        else if(price[i] > price[i - 1])
        {
            sign <- c(sign, 1)
        }
        else
        {
            sign <- c(sign, sign[i - 1])
        }
    }
    return(sign)
}

# Create a function that implements the Lee-Ready Rule.
lee_ready_rule <- function(price)
{
    tick <- tick_test(price)
    sign <- c(1)
    bid <- sapply(tqdata$BID, FUN = as.numeric)
    ask <- sapply(tqdata$OFR, FUN = as.numeric)

    for(i in 2:(length(price)))
    {
        mid <- (bid[i] + ask[i]) * 0.5

        if(price[i] > mid)
        {
            sign <- c(sign, 1)
        }
        else if(price[i] < mid)
        {
            sign <- c(sign, -1)
        }
        else 
        {
            sign <- c(sign, tick[i])
        }
    }
    return(sign)
}

Lee_Ready_Rule_TQ <- lee_ready_rule(price)

Lee_Ready_Rule_Actual <- getTradeDirection(tqdata)

# Check to see if Lee-Ready implementation is the same.
length( which(Lee_Ready_Rule_Actual == Lee_Ready_Rule_TQ ) ) / length(Lee_Ready_Rule_TQ)


# Check to see accuracy of Lee-Ready Rule.
length( which(sign_converted ==  Lee_Ready_Rule_Actual) ) / length(sign_converted)
length( which(sign_converted ==  Lee_Ready_Rule_TQ) ) / length(sign_converted)
```

As such, the Lee-Ready rule is $74.3\%$ accurate in terms of the trade signs it correctly predicted.

